---
title: "1. Fitting a sensor"
output: html_notebook
---

# Aim

This is notebook to show how to:

* Make some demo sensor data with bias and sensitivity drift
* How to make some simple plots of this data
* How to fit a linear line

# Create the data

In this demo we will be looking at a sensor that is experiencing different kinds of errors.

A sensor fresh out of the box, or over time, can start to misbehave.

* Bias - Sensor is constantly reading higher or lower than it should be
* Sensitivity - Sensor over or under-reads a value

In this demo we are going to construct a simple data-frame `df` that has the following columns

* `date` - Date an observation occurs and is measured by a sensor
* `truth` - The truth
* `y_bias` - Reading from a sensor with a nice simple bias issue
* `y_sensitivity ` - Reading from a sensor with a nice simple sensitivity issue
* `y_both` - Reading from a sensor with a nice simple bias and sensitivity issue
* `y_vary` - Reading from a sensor with a nice Simple bias but complex sensitivity issue (gets more sensitive over time)

```{r Make Data, fig.height=8, fig.width=16}
make_data <- function(n=100) {
  t <- seq(n)
  d <- ISOdate(2018,1,1) + hours(t)
  truth <- sin(t * 2 * pi / 24) # 1 signal per day
  m <- 3
  c <- 2
  y_bias <- truth + c # Nice simple bias : truth = y_bias - 2
  y_sensitivity <- m * truth # Nice simple sensitivity  : truth = y_bias / 3
  y_both <- m * truth + c # Nice simple bias and sensitivity  : truth = (y_bias - 2) / 3
  y_vary <- t/24. * truth + c # Simple bias but sensitivity changing over time  : truth = (y_bias - 2) * 24 / t
  df <- data.frame("date" = d,
                   "hour" = t, 
                   "truth" = truth, 
                   "y_bias" = y_bias, 
                   "y_sensitivity" = y_sensitivity, 
                   "y_both" = y_both,
                   "y_vary" = y_vary)
  df
}


df <- make_data(n = 1000)
y_type <- c("y_bias", "y_sensitivity", "y_both", "y_vary")
t_name <- c("slope", "intercept")

df %>% 
  ggplot(aes(x = date, y = truth)) +
  geom_line() +
  xlab(NULL) +
  ylab("Signal (units)") +
  labs(title = "Observed Truth")
```

## Plot the data
```{r Plot Data, fig.height=8, fig.width=8}
df %>% 
  gather(s, val, y_bias, y_sensitivity, y_both, y_vary) %>%
  ggplot(aes(x = date, y = val, color = s)) +
  geom_line() +
  geom_line(aes(x = date, y = truth), color = "black", linetype = 3) +
  facet_grid(s ~ ., scales = "free") +
  xlab(NULL) +
  ylab("Signal (units)") +
  labs(title = "Time series plot") +
  theme(legend.position = "none")

df %>% 
  gather(s, val, y_bias, y_sensitivity, y_both, y_vary) %>%
  ggplot(aes(x = truth, y = val, color = s)) +
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  facet_wrap(s ~ ., scales = "free") + 
  geom_abline() +
  xlab("Truth") +
  ylab("Signal (units)") +
  labs(title = "Scatter plot") +
  theme(legend.position = "none")
```

# How to find the m and c value when fitting data?

* To create a model 
    * Functional model.  In our case we are using a linear model so `lm`
    * A formula of the form `observation ~ sensor`
    * Some data! `df`
    
* To explore the model
    * `summary`
    * `coef`
    * `print`
    
```{r Linear Model Demo, fig.height=8, fig.width=8}
# Make the model
f <- as.formula("truth ~ y_bias")
model <- lm(f, df)

# Explore the model
summary(model)
coef(model)
print(model)

# Use the model
intercept <- model$coefficients[1]
slope <- model$coefficients[2]

result <- sprintf("Formula is truth = y_bias * %.2f + %.2f", slope, intercept)
print(result)
# Recall
# y_bias = truth + 2
# So
# truth = y_bias - 2

df %>% 
  mutate(new_truth = y_bias * slope + intercept) %>%
  gather(s, val, y_bias, new_truth) %>%
  ggplot(aes(x = truth, y = val, color = s)) +
  geom_point(size = 8) +
  stat_smooth(method = "lm") +
  geom_abline() +
  coord_equal() +
  xlab("Truth") +
  ylab("Signal (units)") +
  xlim(-3 , 3) + ylim(-3 , 3)
```

# How to run on many sensors?

## For loop
```{r for loop}
for (yname in y_type) {
  f <- as.formula(sprintf("truth ~ %s", yname))
  model <- lm(f, df)
  intercept <- model$coefficients[1]
  slope <- model$coefficients[2]
  message(sprintf("Formula is truth = %.2f(%s + %.2f)", slope, yname, intercept))
}
```
