---
title: "2. Toy model"
output: html_notebook
---

# Aim

This is notebook to show how to:

* How to run code many times using:
    * for loops
    * foreach and parallel

# For loop
```{r for loop}
yseq <- seq(5)
x <- 0
for (y in yseq) {
  Sys.sleep(2)
  x <- x + 3
  logging::loginfo(sprintf("Working on sequential loop %d, x = %d", y, x))
}

logging::loginfo(x)
```


# Using multiple processors

When you are using multiple processors you need to do a bit of setup.

* How many processors/cores do you have
* What kind of process will you run
* Make a cluster
* Register them
* What variables should all clusters know about?
* Where should error messages go? What libraries should be loaded?
* What happens when you close R?
* Do the work!
* Shut it down

```{r}
# Value to process
yseq <- seq(5)
x <- 0

# Get the number of cores available
numCores <- detectCores() 

# What kind of process?
# (PSOCK and FORK (not available on windows!) are other options)
sock <- "PSOCK" 

# Make a cluser and register it
myCluster <- makeCluster(numCores, type = sock, rscript_args = "--vanilla")
registerDoParallel(myCluster)

# What variables should each cluster know about?
clusterExport(myCluster, "x")

# Where do my errors go?  What do I want to setup?
tmp <- clusterEvalQ(myCluster, 
                    {sink(paste0("../logs/log.toy.", sys.parent(), ".", Sys.getpid(),".txt"));
                     library(tidyverse)})

# What to do if program stopes
on.exit(stopCluster(myCluster))

# How to combine all the end results
mycombinefunc <- function(a,b){a + b}

# Here is the for loop in ||
new_x <- foreach(y = yseq,
                 .combine = "mycombinefunc") %dopar% {
                    Sys.sleep(2)
                    x <- x + 3
                    logging::loginfo(sprintf("Working on || loop %d, x = %d", y, x))
                    x
                  }

logging::loginfo(new_x)

# IMPORTANT Shut it down
stopCluster(myCluster)
```

# Benchmark them against each other
```{r Benchmark}
# for loop
loop_process <- function() {
  yseq <- seq(5)
  x <- 0
  for (y in yseq) {
    Sys.sleep(2)
    x <- x + 3
    logging::loginfo(sprintf("Working on sequential loop %d, x = %d", y, x))
  }
  logging::loginfo(x)
}

# Value to process
p_process <- function() {
  yseq <- seq(5)
  x <- 0
  numCores <- length(yseq) 
  sock <- "PSOCK" 
  myCluster <- makeCluster(length(yseq), 
                           type = sock, 
                           rscript_args = "--vanilla")
  registerDoParallel(myCluster)
  mycombinefunc <- function(a,b){a + b}
  clusterEvalQ(myCluster, sink(paste0("../logs/log.toy.", sys.parent(), ".", Sys.getpid(),".txt")))

  # Note that the .packages, .exports can defined in the foreach function
  new_x <- foreach(y = yseq,
                   .combine = "mycombinefunc") %dopar% {
                       Sys.sleep(2)
                       x <- x + 3
                       logging::loginfo(sprintf("Working on || loop %d, x = %d", y, x))
                       x
                     }
  logging::loginfo(new_x)
  stopCluster(myCluster)
}

# Benchmark 
mbm <- microbenchmark("LOOP" = loop_process(), 
                      "PARALLEL" = p_process(), 
                      times = 1)
print(mbm)
```
